Forcing for Isabelle

### 2018-12-28 LICS/intro_m.tex
Más trabajo relacionado. Menos pegados a fabstracts.

The \emph{Formal Abstract} project~\cite{hales-fabstracts} proposes
that the mechanization of mathematics can advance by formalizing some
main result under the explicit assumption of the material
(definitions, propositions, lemmas) on which it is based. In line with
this advice, we profited that the proof that those axioms hold in
generic extension is independent of the \emph{proofs} of the
``fundamental'' theorems of forcing. Let us remark that the definition
of the forces relation is, by itself, quite demanding; the
formalization of it and of the fundamental theorems of forcing roughly
comprises little less than a half of our full project.

It might be a little surprising the lack of formalizations of forcing
and generic extensions. As far as we know, the development of
\citet{JFR6232} in homotopy type theory for constructing generic
extensions in a sheaf-theoretic setting is the unique mechanization of
forcing. This contrast with the fruitful use of forcing techniques to
extend the Curry-Howard isomorphism to classical axioms
\cite{Miquel:2011:FPT:2058525.2059614,lmcs:1070}. Moreover, the
combination of forcing with intuitionistic type theory
\cite{coquand2010note} gives rise both to positive results (an
algorithm to obtain witnesses of the continuity of definable
functionals \cite{coquand2012computational}) and also negative
(the independence of Markov's principle \cite{lmcs:3859}).

% moreover forcing translations \cite{jaber:hal-01319066}.

>>>>

The \emph{Formal Abstract} project~\cite{hales-fabstracts} proposes
that the formalization of mathematics by writing the statements of
results and the material upon which they are based (definitions,
propositions, lemmas), but ommiting the proofs. We took a less drastic
position: since the proof that those axioms hold in generic extension
is independent of the \emph{proofs} of the ``fundamental'' theorems of
forcing, we assumed these results. Let us remark that the definition
of the forces relation is, by itself, quite demanding; the
formalization of it and of the fundamental theorems of forcing roughly
comprises little less than a half of our full project.

It might be a little surprising the lack of formalizations of forcing
and generic extensions. As far as we know, the development of
\citet{JFR6232} in homotopy type theory for constructing generic
extensions in a sheaf-theoretic setting is the unique mechanization of
forcing. This contrast with the fruitful use of forcing techniques to
extend the Curry-Howard isomorphism to classical axioms
\cite{Miquel:2011:FPT:2058525.2059614,lmcs:1070}. Moreover, the
combination of forcing with intuitionistic type theory
\cite{Coquand:2009:FTT:1807662.1807665,coquand2010note} gives rise
both to positive results (an algorithm to obtain witnesses of the
continuity of definable functionals \cite{coquand2012computational})
and also negative (the independence of Markov's principle
\cite{lmcs:3859}). In the same strand of forcing from the point of
view of proof theory \cite{avigad_2004} are the conservative
extensions of CoC with forcing conditions
\cite{jaber:hal-01319066,Jaber:2012:ETT:2358958.2359524}.
